SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-07 docs/questioning.tex
some changes

Application attacks that involve {\sc dsl}s constitute an important
subset of the code injection problem, as {\sc dsl}s like {\sc sql} and
{\sc xml} play an significant role in the development of either web or
mobile applications. For example, many applications have interfaces
where a user enters input to interact with the application's data,
thereby interacting with the underlying Relational Database Management
System ({\sc rdbms}). This input can become part of an {\sc sql} query
and executed on the target {\sc rdbms}. A code injection attack that
exploits the vulnerabilities of these interfaces by taking advantage
of input validation issues like incorrectly passed parameters or
incorrect type handling, is called an {\sc sql} injection
attack~\cite{CERT02,MS09,HVO06,SW06}. By using similar techniques
malicious users can perform other exploits based on {\sc dsl}s,
like {\sc xp}ath~\cite{SW06,CDL07,MKS09}, {\sc xml}~\cite{MSM13}
and No{\sc sql}~\cite{SMS13}. With this kind of attacks,
a malicious user can view sensitive information,
destroy or modify protected data, or even crash the entire
application. {\sc html} is another {\sc dsl} that can be used
for malicious purposes when an application does not properly
handle user supplied data. Based on this vulnerability
an attacker can supply valid {\sc html},
typically via a parameter value, and inject their own
content into the page. {\sc html} injection is mainly associated
with {\sc xss} defects.

@@

Training methods that detect {\sc dsl}-driven injection attacks
generate and store valid code statements (i.e. {\sc sql} or {\sc
  xp}ath queries) in various forms, and detect attacks as outliers
from the set of valid code statements. An early approach, {\it {\sc
    didafit}}~\cite{LLW02} detects {\sc sql} injection (Locus
$\Omega$) attacks by recording all database transactions. Subsequent
refinements by Valeur et al.~\cite{VMV05} tagged each transaction with
the corresponding application as an extension of their anomaly
detection framework called {\it libAnomaly}. {\it {\sc
    sd}river}~\cite{MS09,MKS09,MKLS11} is a signature-based mechanism
that prevents {\sc sql} and {\sc xp}ath injection attacks. The
signatures generated during a training phase are based on features
that can depend either on the code statement or on its execution
environment (i.e. the stack trace). Then, at runtime, the mechanism
checks all statements for compliance and can block code statements
containing injected elements (Locus {\sc z}). {\sc
  amnesia}~\cite{HO05,HO06,HO05b} is a tool that also detects {\sc
  sql} injection attacks (locus {\sc z}) by associating a query model
with the location of every {\sc sql} statement within the application.
Then, at runtime, it monitors the application's execution to detect
when {\sc sql} statements diverge from the expected model.

@@

Defenses based on training include some mechanisms that can be easily
circumvented. For example, {\sc didafit} records all database
transactions without tagging each transaction with the
corresponding call-site. In this case a false negative result
is prominent. In particular, consider an application that will send the
password for a forgetful user via email by executing the following
query: \bgroup \lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
A traditional {\sc sql} injection attack protection system like {\sc didafit},
could be easily escaped since a malicious user, could obtain the
administrator's password by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent
%% PL I do not understand what is specific to training here; most
%% classic SQL injection attacks work in this way.
However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
Furthermore, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect a return-to-JavaScript
attack~\ref{sec:policy}.
%% PL Like the return-to-lib attack?

@@

    \item[3] For every overhead we show if the overhead is applied to the server ({\sc s}) or the client ({\sc c}). 

@@

% Cite~\cite{HNSHS12}.

>>>>

Application attacks that involve {\sc dsl}s constitute an important
subset of the code injection problem, as {\sc dsl}s like {\sc sql} and
{\sc xml} play an significant role in the development of either web or
mobile applications. For example, many applications have interfaces
where a user enters input to interact with the application's data,
thereby interacting with the underlying Relational Database Management
System ({\sc rdbms}). This input can become part of an {\sc sql} query
and executed on the target {\sc rdbms}. A code injection attack that
exploits the vulnerabilities of these interfaces by taking advantage
of input validation issues like incorrectly passed parameters or
incorrect type handling, is called an {\sc sql} injection
attack~\cite{CERT02,MS09,HVO06,SW06}. By using similar techniques
malicious users can perform other exploits based on {\sc dsl}s,
like {\sc xp}ath~\cite{SW06,CDL07,MKS09}, {\sc xml}~\cite{MSM13}
and No{\sc sql}~\cite{SMS13}. With this kind of attacks,
a malicious user can view sensitive information,
destroy or modify protected data, or even crash the entire
application. {\sc html} is another {\sc dsl} that can be used
for malicious purposes when an application does not properly
handle user supplied data. Based on this vulnerability
an attacker can supply valid {\sc html},
typically via a parameter value, and inject their own
content into the page. {\sc html} injection is mainly associated
with {\sc xss} defects~\cite{HNSHS12}.

@@

Training methods that detect {\sc dsl}-driven injection attacks
generate and store valid code statements (i.e. {\sc sql} or {\sc
  xp}ath queries) in various forms, and detect attacks as outliers
from the set of valid code statements. An early approach, {\it {\sc
    didafit}}~\cite{LLW02} detects {\sc sql} injection (Locus
$\Omega$) attacks by recording all database transactions
(stripped from user input). Subsequent
refinements by Valeur et al.~\cite{VMV05} tagged each transaction with
the corresponding application as an extension of their anomaly
detection framework called {\it libAnomaly}. {\it {\sc
    sd}river}~\cite{MS09,MKS09,MKLS11} is a signature-based mechanism
that prevents {\sc sql} and {\sc xp}ath injection attacks. The
signatures generated during a training phase are based on features
that can depend either on the code statement or on its execution
environment (i.e. the stack trace). Then, at runtime, the mechanism
checks all statements for compliance and can block code statements
containing injected elements (Locus {\sc z}). {\sc
  amnesia}~\cite{HO05,HO06,HO05b} is a tool that also detects {\sc
  sql} injection attacks (locus {\sc z}) by associating a query model
with the location of every {\sc sql} statement within the application.
Then, at runtime, it monitors the application's execution to detect
when {\sc sql} statements diverge from the expected model.

@@

Defenses based on training include some mechanisms that can be easily
circumvented. For example, {\sc didafit} records all database
transactions without tagging each transaction with the
corresponding call-site. In this case a false negative result
is prominent. For instance, consider an application that will send the
password for a forgetful user via email by executing the following
query: \bgroup \lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
In this case {\sc didafit}, could be easily escaped
since a malicious user, could obtain the administrator's password
by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent
%% PL I do not understand what is specific to training here; most
%% classic SQL injection attacks work in this way.
However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
Furthermore, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect a return-to-JavaScript
attack~\ref{APKLM10}.
%% PL Like the return-to-lib attack?

@@

    \item[3] Regarding the computational overhead we show if it is applied to the server ({\sc s}) or the client ({\sc c}). 

@@
